Verification of OpenSSL's SHA3 Keccak function using the Software Analysis Workbench

### 2021-09-10 paper/abstract.tex
update abstract and figure

\begin{abstract}
The Keccak function in the SHA-3 NIST standard is responsible for the major computations of all SHA-3 hashes and Shake functions.
These functions are a key piece in modern secure hash functions. 
This work extends and improves on the methodology used in the verification of Amazon’s s2n HMAC and OpenSSL’s SHA2-256 inner hash function. 
It uses the Software Analysis Workbench (SAW) to prove out functional equivalence between OpenSSL’s SHA-3 Keccak function and a verified implementation of SHA-3 in the domain-specific language Cryptol. 
SAW creates contracts (called overrides), a functions input to output relationship, which allows for a modular approach to writing proofs. 
This work leverages SAW by creating overrides to block out computation and creating a compositional proof that remains within the capacity of an SMT solver. 
This work amends Cryptol through utility functions to handle looping which greatly simplifies the visual inspection the Cryptol implementation to provide strong assurance of correctness. 
The results of this work indicate that this methodology works with modern cryptographic specifications and that a modular approach in cryptographic algorithm design greatly simplifies the verification process.
\end{abstract}

>>>>

\begin{abstract}
The Keccak function in the SHA-3 NIST standard is responsible for the major computations of all SHA-3 hash constructs.
These constructs form the basis for modern secure hashing.
This work extends and improves the methodology used to verify Amazon’s s2n HMAC and OpenSSL’s SHA2-256 inner hash function. 
It uses the Software Analysis Workbench (SAW) to prove out functional equivalence between OpenSSL’s SHA-3 Keccak function and a verified implementation of SHA-3 in Cryptol, a domain-specific language for cryptography. 
SAW allows for a modular approach to writing proofs through the use of contracts, which define a functions input to output relationship.
Each contract, also called an override, is an obligation in the full proof.
This work leverages SAW by creating overrides to decompose computation and create a compositional proof that is within the capacity of an SMT solver. 
Visual inspection between Cryptol and the NIST standard is simplified through the creation of Cryptol utility functions.
These utility functions wrap the complex list comprehensions required for looping in Cryptol.
The results of this work indicates that a modular approach in cryptographic algorithm design greatly simplifies the verification process, as exemplified by SHA-3.
\end{abstract}

### 2021-09-10 paper/introduction.tex
Add limitations to the introduction
update abstract and figure

  The reliability of this correctness check is often weakened because of the difficulty of performing that visual review.  
  Many implementations of the SHA algorithms go through major rewrites and optimizations which make visual inspection of the code nearly impossible. (Not sure if that’s too strong a claim).
  This work bridges that gap between optimized c code, and the NIST standard.

\noindent \textbf{Contributions}:
This paper details the following contributions.
\begin{compactitem}
  \item  A Cryptol specification for \texttt{SHA3} that directly corresponds to the \texttt{FIPS} 202 specification and passes the test vectors provided by the same specification.
  \item A Cryptol specification based on the OpenSSL implementation of \texttt{SHA3} that uses a reordered state vector and precomputed tables in Keccak for performance gains and passes test vectors
  \item A \texttt{SAW} proof that the \texttt{FIPS} 202 specification is equivalent to the OpenSSL specification
  \item A \texttt{SAW} proof that the OpenSSL Keccak specification is equivalent to the OpenSSL C-implementation of the same function for all possible state vector input
  \item Early evidence that modular cryptographic specifications such as \texttt{SHA3} lend themselves to efficient \texttt{SAW} verification against C implementations as opposed to flattened specifications such as \texttt{SHA2} 
\end{compactitem}

@@

>>>>

  The reliability of this correctness check is often weakened because of the difficulty of performing that visual review.  
  Many implementations of the SHA algorithms go through major rewrites and optimizations which make visual inspection of the code nearly impossible. (Not sure if that’s too strong a claim).
  This work bridges that gap between optimized c code, and the NIST standard.
  \input{proofFigure.tex}
\noindent \textbf{Contributions}:
This paper details the following contributions.
\begin{compactitem}
  \item  A Cryptol specification for \texttt{SHA3} that directly corresponds to the \texttt{FIPS} 202 specification and passes the test vectors provided by the same specification.
  \item A Cryptol specification based on the OpenSSL implementation of \texttt{SHA3} that uses a reordered state vector and precomputed tables in Keccak for performance gains and passes test vectors
  \item A \texttt{SAW} proof that the \texttt{FIPS} 202 specification is equivalent to the OpenSSL specification
  \item A \texttt{SAW} proof that the OpenSSL Keccak specification is equivalent to the OpenSSL C-implementation of the same function for all possible state vector input
  \item Early evidence that modular cryptographic specifications such as \texttt{SHA3} lend themselves to efficient \texttt{SAW} verification against C implementations as opposed to flattened specifications such as \texttt{SHA2} 
\end{compactitem}

\noindent \textbf{Limitations}
\begin{compactitem}
  \item Visual inspection and limited test vectors for (1) and (2)
  \item Only provide proof certificates for message sizes of X to Y for digest sizes of … for (3)
  \item Visual inspection only for anything above Keccak for (3) in OpenSSLs C-code
\end{compactitem}

@@

Because of how NIST defines the standard, the keccak function is used for each digest size of the SHA3 algorithm, with the only significant differences being in the padding/sponging parts of the algorithm, which falls outside this work's scope.
This means that the proof of correctness for this function extends usefulness beyond just the SHA3-256 level which we focused on.
In our repository, we wrote the Cryptol specification to be able to run any of the SHA3 family of algorithms using the same keccak.
A simple extension of this work would be to take that already created and verified Cryptol specification, and prove it up against OpenSSL's slightly modified keccak versions.
